---
title: "RLabFirstProj"
author: "mahdiyeh ebrahimy"
date: "2024-11-21"
output:
  html_document:
    df_print: paged
---
<h1 style="font-size: 25px;">Exploring the Salaries Dataset</h1>

```{r}
my.id <- 20 + 81 
set.seed(my.id)
my.id
#install.packages("carData")
data <- carData::Salaries
head(data,10)
```
<h2 style="font-size: 20px;">Descriptive Statistics and Visualization of Sampled yrs.service Data</h2>
```{r}
x <- sample(data$yrs.service, my.id, replace = F)
mean(x)
median(x)
sd(x)
var(x)
range <- max(x)-min(x)
range
summary(x)
```
The first quartile (7.00) indicates that 25% of the data is less than or equal to 7.00.
The median (16.00) indicates that 50% of the data is less than or equal to 16.00.
The third quartile (23.00) indicates that 75% of the data is less than or equal to 23.00.
```{r}
#install.packages("e1071")
e1071::skewness(x,type = 3)
e1071::kurtosis(x,type = 3)
hist(x,col="lightblue")
```

<div dir="rtl">
  <h2 style="font-size: 20px;">مقدار چولگی مثبت است پس یعنی داده ها چولگی به سمت راست دارند این چیزی است که هیستوگرام هم به ما تقریبا نشان می دهد همچنین داده ها کشیدگی کمتری از داده های نرمال(0) دارند</h2>
</div>
<h1 style="font-size: 25px;">Job Rank Counts and Salary histogram</h1>
```{r}
my.id <- 20 + 81 
set.seed(my.id)
sampled_data <- data[sample(nrow(data), my.id, replace = F), ]
x <- sampled_data$rank
barplot(table(x),
        col = "lightblue",
        names.arg = c("Assistant Professor", "Associate Professor", "Professor"),
        main = "Job Rank Counts",  
        xlab = "Job Rank",             
        ylab = "Frequency")           
y <-  sampled_data$salary
hist(y, 
     col = "lightblue", 
     main = "Salary Histogram", 
     xlab = "Salary",     
     ylab = "Frequency")   
boxplot(y,
        col = "lightblue",
        main = "Salary boxPlot")
x_copy <- as.numeric(x)
plot(x_copy, y, 
     xlab = "Rank", 
     ylab = "Salary", 
     main = "Rank and Salary",
     pch = 19,   
     col = "lightblue",
     xaxt = 'n') 
     axis(1, at = 1:3, labels = c("Assistant Professor", "Associate            Professor", "Professor"))
```
<h1 style="font-size: 25px;">random generating from normal distribution</h1>
```{r}
my.id <- 20 + 81 
set.seed(my.id)
x <- round(rnorm(100,50,10),0)
x_6 <- x[x %% 6 == 0]
print(x_6)
```
<h1 style="font-size: 25px;">function</h1>
```{r}
f <- function(x){
  n <- length(x)
  return(n/sum(1/x))
}
f(data$salary)
```
<h1 style="font-size: 25px;">Integrate</h1>
```{r}
f1 <- function(x) {
  1 + x^2
}
#install.packages("pracma")
pracma::integral(f1,1/2,3, reltol = 1e-10)
f2 <- function(x, y) {
  x + y^2 - 1
}
pracma::integral2(f2, 3, 5, 0, 1, reltol = 1e-10)
```
<h1 style="font-size: 25px;">matrix</h1>
```{r}
A <- matrix(c(1, 2, 4, 6, 1, 3, 9, 5, 3), nrow = 3, byrow = TRUE)
det(A)
if (det(A)!= 0) {
  solve(A)
} else {
  print("Matrix is singular")
}
diag(A) = c(0,0,0)
print(A)
```
